---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# MATSS `r packageVersion("MATSS")`

[![Build Status](https://travis-ci.org/weecology/MATSS.svg?branch=master)](https://travis-ci.org/weecology/MATSS)
[![License](https://img.shields.io/badge/license-MIT-blue.svg)](https://raw.githubusercontent.com/weecology/MATSS/master/LICENSE)
[![Coverage
status](https://codecov.io/gh/weecology/MATSS/branch/master/graph/badge.svg)](https://codecov.io/github/weecology/MATSS?branch=master)



## Overview
The **`MATSS`** package is intended to support Macroecological Analysis of Time Series Structure. We provide functions to:

* gather ecological time series datasets
* perform basic processing and summaries of those datasets
* build an analytical pipeline to conduct macroecological analyses on those datasets
* create template reports for collating results and produce syntheses

## Contributing

For more information about contributing code, datasets, or analyses, please check out the [Contributing Guide](CONTRIBUTING.md).

## Installation

You can install **`MATSS`** from github with:

```{r, eval = FALSE}
# install.packages("remotes")
remotes::install_github("weecology/MATSS", build_opts = c("--no-resave-data", "--no-manual"))
```

This package relies on the development version of the `rdataretriever` package
to install datasets. Installation of this package takes a few extra steps
because it runs a Python package behind the scenes. Follow the installation
instructions on the [`rdataretriever` README](https://github.com/ropensci/rdataretriever).

## Datasets

```{r setup packages, include = FALSE}
library(MATSS)
library(drake)
```

```{r compute info on datasets, include = FALSE, cache = TRUE}
datasets <- build_datasets_plan(include_retriever_data = TRUE, 
                                include_bbs_data = TRUE, 
                                include_gpdd_data = TRUE, 
                                include_biotime_data = TRUE)

count_ts <- function(data) {NCOL(data$abundance)}

analyses <- drake_plan(num_ts = target(count_ts(data), 
                                       transform = map(data = !!rlang::syms(datasets$target))), 
                       total = target(sum(num_ts), 
                                      transform = combine(num_ts)))

plan <- rbind(datasets, analyses)
make(plan)
```

**`MATSS`** pulls data from a variety of sources, including:

* `r NROW(build_datasets_plan(include_retriever_data = TRUE))` individual datasets that we've added,
* the North American Breeding Bird Survey database (spanning `r NROW(build_bbs_datasets_plan())` separate datasets),
* the Global Population Dynamics Database (spanning `r NROW(build_gpdd_datasets_plan())` separate datasets),
* and the BioTime database (spanning `r NROW(build_biotime_datasets_plan())` separate datasets).

Combined, there are **`r readd(total)`** individual time series across all of these data sources.

## Getting Started

To get started with the data or analysis templates, we recommend you take a look at our [Getting Started vignette](https://weecology.github.io/MATSS/articles/MATSS.html) for more details about how to interface with the datasets, use Drake to create workflows, and create research compendia.

If you have the **`MATSS`** package installed, you can also view the vignette from within R:

```{r, eval = FALSE}
vignette("MATSS")
```

## Example Use Cases

Here are some examples of using **`MATSS`** to create research compendia:

* [MATSS-LDATS](https://github.com/weecology/MATSS-LDATS) applies the [**`LDATS`**](https://github.com/weecology/LDATS) package to investigate changepoints in community dynamics across the datasets in **`MATSS`**
* [MATSS-Forecasting](https://github.com/weecology/MATSS-forecasting) investigates which properties are associated with the predictability of population time series across the datasets in **`MATSS`**
