---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# MATSS `r packageVersion("MATSS")`

<!-- badges: start -->
[![Build Status](https://travis-ci.org/weecology/MATSS.svg?branch=master)](https://travis-ci.org/weecology/MATSS)
[![License](https://img.shields.io/badge/license-MIT-blue.svg)](https://raw.githubusercontent.com/weecology/MATSS/master/LICENSE)
[![Coverage
status](https://codecov.io/gh/weecology/MATSS/branch/master/graph/badge.svg)](https://codecov.io/github/weecology/MATSS?branch=master)
[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.3333008.svg)](https://doi.org/10.5281/zenodo.3333008)
<!-- badges: end -->

## Overview
The **`MATSS`** package is intended to support Macroecological Analysis of Time Series Structure. We provide functions to:

* gather ecological time series datasets
* perform basic processing and summaries of those datasets
* build an analytical pipeline to conduct macroecological analyses on those datasets
* create template reports for collating results and produce syntheses

## Contributing

For more information about contributing code, datasets, or analyses, please check out the [Contributing Guide](CONTRIBUTING.md).

## Installation

You can install **`MATSS`** from github with:

```{r, eval = FALSE}
# install.packages("remotes")
remotes::install_github("weecology/MATSS", build_opts = c("--no-resave-data", "--no-manual"))
```

This package relies on the development version of the `rdataretriever` package
to install datasets. Installation of this package takes a few extra steps
because it runs a Python package behind the scenes. Follow the installation
instructions on the [`rdataretriever` README](https://github.com/ropensci/rdataretriever).

## Datasets

```{r setup packages, include = FALSE}
library(MATSS)
library(drake)
```

```{r compute info on datasets, include = FALSE, cache = TRUE}
datasets <- build_datasets_plan(include_retriever_data = TRUE, 
                                include_bbs_data = TRUE, 
                                include_gpdd_data = TRUE, 
                                include_biotime_data = TRUE)

count_ts <- function(data) {NCOL(data$abundance)}

analyses <- drake_plan(num_ts = target(count_ts(data), 
                                       transform = map(data = !!rlang::syms(datasets$target))), 
                       total = target(sum(num_ts), 
                                      transform = combine(num_ts)))

plan <- rbind(datasets, analyses)
make(plan)
```

**`MATSS`** pulls data from a variety of sources, including:

* `r NROW(build_datasets_plan(include_retriever_data = TRUE))` individual datasets that we've added,
* the North American Breeding Bird Survey database (spanning `r NROW(build_bbs_datasets_plan())` separate datasets),
* the Global Population Dynamics Database (spanning `r NROW(build_gpdd_datasets_plan())` separate datasets),
* and the BioTime database (spanning `r NROW(build_biotime_datasets_plan())` separate datasets).

Combined, there are **`r readd(total)`** individual time series across all of these data sources.

## Getting Started

To get started with the data or analysis templates, we recommend you take a look at our [Getting Started vignette](https://weecology.github.io/MATSS/articles/MATSS.html) for more details about how to interface with the datasets, use Drake to create workflows, and create research compendia.

If you have the **`MATSS`** package installed, you can also view the vignette from within R:

```{r, eval = FALSE}
vignette("MATSS")
```

## Example Use Cases

Here are some examples of using **`MATSS`** to create research compendia:

* [MATSS-LDATS](https://github.com/weecology/MATSS-LDATS) applies the [**`LDATS`**](https://github.com/weecology/LDATS) package to investigate changepoints in community dynamics across the datasets in **`MATSS`**
* [MATSS-Forecasting](https://github.com/weecology/MATSS-forecasting) investigates which properties are associated with the predictability of population time series across the datasets in **`MATSS`**

## Acknowledgments

We thank Erica Christensen and Joan Meiners for their contributions and input on early prototypes of this project. This project would not be possible without the support of Henry Senyondo and the [retriever](https://www.data-retriever.org/) team. 

Package development is supported through various funding sources: including the [Gordon and Betty Moore Foundation’s Data-Driven Discovery Initiative](http://www.moore.org/programs/science/data-driven-discovery), [Grant GBMF4563](http://www.moore.org/grants/list/GBMF4563) to E. P. White (supporting also the time of J. Simonis and H. Ye), the [National Science Foundation](http://nsf.gov/), [Grant DEB-1622425](https://www.nsf.gov/awardsearch/showAward?AWD_ID=1622425) to S. K. M. Ernest, and a [National Science Foundation Graduate Research Fellowship](https://www.nsfgrfp.org/) 
(No. [DGE-1315138](https://www.nsf.gov/awardsearch/showAward?AWD_ID=1315138) 
and [DGE-1842473](https://www.nsf.gov/awardsearch/showAward?AWD_ID=1842473)) to R. Diaz.
