---
title: "Data Formats"
author: 
- Ellen K. Bledsoe
- Hao Ye
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## This is data should be formatted for the `run_LDA` functions:
a data.frame or tibble; each row is an observation (e.g. in time or space), and each column is a variable. Here, the common usage is for each column to be a species or taxon, and each row to be an observed sample. In the original specification for LDA, each row is a document, and each column is a word, with the entries being the counts of the words in each document.

```{r example crosstab data, echo = FALSE}
data <- data.frame("Red Spotted Dragon" = c(2,6,0,5,4,4),
                   "Green Striped Dragon" = c(6,0,4,1,9,7),
                   "Blue Eyes White Dragon" = c(0,0,0,1,0,0))
knitr::kable(data) 
```