---
title: "MATSS: Dataset Summary"
author: 
- Hao Ye
date: "`r Sys.Date()`"
output:
  rmarkdown::html_document:
    self_contained: yes
fontsize: 11pt
documentclass: article
vignette: >
  %\VignetteIndexEntry{MATSS: Dataset Summary}
  %\VignetteEngine{knitr::rmarkdown_notangle}
  %\VignetteEncoding{UTF-8}
---
  
```{r setup, include = FALSE}
IS_LOCAL <- identical(Sys.getenv("LOCAL"), "true")

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  purl = IS_LOCAL
)
```

```{r setup packages}
library(dplyr)
library(tidyr)
library(purrr)
library(MATSS)
```

# Overview

The basic info we want for each dataset is:

* taxon data (species-level if appropriate)
* sampling start and end times
* sampling frequency
* geographical location
* life history characteristics of the organisms
* temperature at sampling location

For the latter two, it's probably easier to rely on external packages to retrieve the info, so the summaries will first focus on the remaining entries.

# BioTime summaries

Define a function for how to generate the summary for BioTime datasets. Note that we use a `tibble` structure, to enable easy assembly across datasets.

```{r define summary for Biotime data}
summarize_biotime_data <- function(dataset_id = 10, ...)
{
    dat <- get_biotime_data(dataset_id = dataset_id, 
                            ...)
    
    summ_info <- tibble(dataset_id = dataset_id)
        
    # determine species info
    summ_info$species_table <- I(list(dat$metadata$species_table))
    
    # determine start and end time
    if (dat$metadata$is_annual_sampling)
    {
        var <- "year"
    } else {
        var <- "date"
    }
    summ_info$start_time <- as.character(min(dat$covariates[[var]], na.rm = TRUE))
    summ_info$end_time <- as.character(max(dat$covariates[[var]], na.rm = TRUE))
    
    # determine sampling frequency
    if (dat$metadata$is_annual_sampling) # missing month data in raw values
    {
        summ_info$sampling_frequency <- "annual (or less)"
    } else if (length(unique(dat$covariates$year)) >= NROW(dat$covariates)) {
        summ_info$sampling_frquency <- "annual (or less)"
    } else {
        summ_info$sampling_frequency <- "subannual"
    }
    
    # determine spatial extent
    summ_info$number_locations <- dat$metadata$number_lat_long
    summ_info$lat <- dat$metadata$cent_lat
    summ_info$long <- dat$metadata$cent_long
    summ_info$lat_min <- min(dat$covariates$latitude, na.rm = TRUE)
    summ_info$lat_max <- max(dat$covariates$latitude, na.rm = TRUE)
    summ_info$long_min <- min(dat$covariates$longitude, na.rm = TRUE)
    summ_info$long_max <- max(dat$covariates$longitude, na.rm = TRUE)
    return(summ_info)
}
```

## Results

Now repeat for all the datasets within BioTime

```{r biotime summaries, eval = IS_LOCAL}
biotime_dataset_ids <- get_biotime_dataset_ids()

biotime_summaries <- map_dfr(biotime_dataset_ids, summarize_biotime_data)
```

Extract out just the dataset-level info:

```{r dataset info, eval = IS_LOCAL}
dataset_info <- biotime_summaries %>% 
    select(-species_table)

knitr::kable(dataset_info)
```

Extract out just the taxa:

```{r species info, eval = IS_LOCAL}
species_info <- biotime_summaries %>%
    unnest() %>%
    select(dataset_id, genus, species, genus_species)

# knitr::kable(species_info)
```

## Search for duplicates
