---
title: "Data Formats"
author: 
- Ellen K. Bledsoe
- Hao Ye
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Data Structure

The universal data structure we're going to use is:

* a list with at least 1 element:
  - a data.frame or tibble, named `abundance` (required)
  - a data.frame or tibble, named `covariates` (optional)
  - a list, named `metadata` (optional)
  
If both `abundance` and `covariates` are present in the list, then the two data.frames must have the same number of rows.

### abundance

In the `abundance` data.frame:

* each row is an observation (e.g. in time or space)
* each column is a variable

Here, the common usage is for each column to be a species or taxon, and each row to be an observed sample. In other words, each column is a time series, with the rows sorted such that time advances down (higher row indices correspond to later times).

### covariates

In the `covariates` data.frame:

* each row is an observation (e.g. in time or space)
* each column is a variable

The number of rows should match that of `abundance`, and rows of `covariates` should line up with `abundance` (either sampled simultaneously or concurrently). Common covariates are date and time, temperature, treatments, etc.

## Example Data

Here is an example of a correctly formatted dataset with covariates:

```{r example crosstab data, echo = FALSE}
abundance <- data.frame("Red Spotted Dragon" = c(2,6,0,5,4,4),
                        "Green Striped Dragon" = c(6,0,4,1,9,7),
                        "Blue Eyes White Dragon" = c(0,0,0,1,0,0))

num_rows <- NROW(abundance)

covariates <- data.frame("date" = rev(seq.Date(from = Sys.Date(), by = "-1 year", length.out = num_rows)), 
                         "precipitation" = rpois(num_rows, lambda = 10))

dat <- mget(c("abundance", "covariates"))

str(dat)
knitr::kable(dat)
```

## Checking Data

We also provide a function for checking whether the data is formatted correctly:

```{r check dat}
library(MATSS)
check_data_format(dat)
```

