---
title: "Installing MATSS on the HiPerGator"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{hipergator-install}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

1. Log onto HiPerGator

    ```bash
    ssh username@hpg2.rc.ufl.edu
    ```

2. Follow the [installation instructions for Python on HiPerGator](https://github.com/weecology/lab-wiki/wiki/Programming:-HiPerGator-Intro-Guide#installing-python-packages). No need to actually install packages yet.

3. Create a file named .Rprofile to allow for local installs

    ```bash
    echo '.libPaths(c("/home/USERNAME/R_libs", .libPaths()))' >> .Rprofile
    ```

4. Load and open the R module (this loads R for use)

    ```bash
    ml R
    R
    ```

5. In `R` install the core Python retriever (yes, this happens in `R`, which is kind of magical).


    ```{r, eval = FALSE}
    install.packages('reticulate')
    library(reticulate)
    conda_install('r-reticulate', 'retriever')
    devtools::install_github('ropensci/rdataretriever')
    ```

6. Restart R

7. Install the `rdataretriever`

    ```{r, eval = FALSE}
    devtools::install_github('ropensci/rdataretriever')
    library(rdataretriever)
    ```

8. Install `MATSS`

    ```{r, eval = FALSE}
    devtools::install_github("weecology/MATSS")
    ```

9. To run a `MATSS` pipeline on HiPerGator you will need to setup `drake` to use
slurm for parallel scheduling. See https://github.com/weecology/MATSS-LDATS/ for
an example of how to do this.