---
title: "Installing MATSS on the HiPerGator"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{hipergator-install}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  eval = FALSE
)
```

0. If you need a HiPerGator account, follow the instructions on the lab-wiki: https://github.com/weecology/lab-wiki/wiki/Programming:-HiPerGator-Intro-Guide

1. Log onto HiPerGator (substituting your username for <USERNAME>).

```{bash}
    ssh <USERNAME>@hpg2.rc.ufl.edu
```

2. Follow the [installation instructions for Python on HiPerGator](https://github.com/weecology/lab-wiki/wiki/Programming:-HiPerGator-Intro-Guide#installing-python-packages). No need to actually install packages yet.

3. Create a file named .Rprofile to allow for local installs (substituting your 
  username for <USERNAME>):

```{bash}
    echo '.libPaths(c("/home/<USERNAME>/R_libs", .libPaths()))' >> .Rprofile
```

4. Make sure that you have the specified folder for storing R packages (substituting your 
  username for <USERNAME>):

```{bash}
   mkdir /home/<USERNAME>/R_libs
```

5. Load the R module (which makes R available in your HiPerGator session), and then open R:

```{bash}
    ml R
    R
```

6. In `R`, install the core Python retriever (yes, this happens in `R`, which is kind of magical).

```{r}
    install.packages('reticulate')
    library(reticulate)
    conda_install('r-reticulate', 'retriever')
```

If you get an error that the conda environment `'r-reticulate'` does not exist, you may need to create it first. Run the following lines and then retry the `conda_install('r-reticulate', 'retriever')` command:

```{r}
    conda_create('r-reticulate')
```

7. Restart R

8. Install the `rdataretriever`

```{r}
    remotes::install_github('ropensci/rdataretriever')
    library(rdataretriever)
```

If you get an error about "HTTP error 404" and "Rate limit remaing", you may need to setup a GITHUB_PAT key on HiPerGator - https://happygitwithr.com/github-pat.html

9. Install `MATSS`

```{r}
    remotes::install_github("weecology/MATSS")
```

10. Try installing a dataset from retriever:
```{r}
    MATSS::install_retriever_data("veg-plots-sdl", force_install = TRUE)
```
    
If you receive an error about the number of arguments, e.g.
```
    Error in py_call_impl(callable, dots$args, dots$keywords) : 
      TypeError: install_csv() takes from 1 to 4 positional arguments but 5 were given
```
then you will need a newer version of retriever than was previously installed.

Exit R to return to the command line, and then use the following commands to install retriever from github:

```{bash}
    conda activate r-reticulate
    conda install git pip
    pip install git+git://github.com/weecology/retriever@master
```

Then, retry the command at the beginning of this step in R.

11. To run a `MATSS` pipeline on HiPerGator you will need to setup `drake` to use
slurm for parallel scheduling. See https://github.com/weecology/MATSS-LDATS/ for
an example of how to do this.