---
title: "MATSS: Getting Started"
author: 
- Hao Ye
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Overview

`MATSS` is a package for conducting Macroecological Analyses of Time Series Structure. In other words, we have designed it with researchers in mind, as a tool for getting started quickly.

The goals of the package are to make it as easy as possible to:
- obtain a hoard of ecological time series data, processed into a common [data format](data-formats.html)
- build an analysis pipeline, using a mixture of functions from the `drake` package and functions that we provide which do some of the background lifting.

## Installation

You can install the `MATSS` package from github with:

``` r
# install.packages("devtools")
devtools::install_github("weecology/MATSS")
```

And load the package in the typical fashion:

```{r}
library(MATSS)
```

# Data

## Packaged datasets

Several datasets are included with this package - these can be loaded individually using specific functions.

```{r, eval = FALSE}
get_jornada_data()
get_maizuru_data()
get_sgs_data()
```

## Downloadable datasets

Other datasets require downloading. To facilitate this, we include functions to help configure a specific location on disk. To check your current setting:

```{r}
get_default_data_path()
```

and to configure this setting:

```{r, eval = FALSE}
use_default_data_path()
```

# Drake Plans

## 


## 


## 





