---
title: "MATSS: Getting Started"
author: 
- Hao Ye
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
%\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  ---
  
  ```{r setup, include = FALSE}
knitr::opts_chunk$set(
collapse = TRUE,
comment = "#>"
)
```

# Overview

`MATSS` is a package for conducting Macroecological Analyses of Time Series Structure. In other words, we have designed it with researchers in mind, as a tool for getting started quickly.

The goals of the package are to make it as easy as possible to:

- obtain a hoard of ecological time series data, processed into a common [data format](data-formats.html)
- build an analysis pipeline, using a mixture of functions from the `drake` package and functions that we provide which do some of the background lifting.

## Installation

You can install the `MATSS` package from github with:

``` r
# install.packages("devtools")
devtools::install_github("weecology/MATSS")
```

And load the package in the typical fashion:

```{r, message = FALSE}
library(MATSS)
```

# Data

## Packaged datasets

Several datasets are included with this package - these can be loaded individually using specific functions.

```{r, eval = FALSE}
get_jornada_data()
get_maizuru_data()
get_sgs_data()
```

## Downloadable datasets

Other datasets require downloading. To facilitate this, we include functions to help configure a specific location on disk. To check your current setting:

```{r, eval = FALSE}
get_default_data_path()
```

and to configure this setting:

```{r, eval = FALSE}
use_default_data_path("<path>")
```

# Working with Drake

For the most part `MATSS` provides only a light wrapper for the functions in `drake`, so it can be helpful to know about how `drake` plans work if you are going to do analyses using `MATSS`. Note that using `drake` plans is not strictly necessary, as you can use the `MATSS` functions in whatever workflow system you desire.

## Basic Workflow

The basic workflow of using `drake` plans in R is:
* run R code that produces `drake` plans
* run R code that takes a `drake` plan and executes it

### Example

```{r, warning = FALSE}
library(drake)
library(dplyr)

# define the plan
plan <- drake_plan(data_1 = mtcars, 
                   data_2 = iris, 
                   my_model = lm(mpg ~ disp, data = data_1), 
                   my_summary = data_2 %>%
                       group_by(Species) %>%
                       summarize_all(mean))

# run the plan
make(plan)

# check resulting objects
readd(my_model)
readd(my_summary)
```

### Making Drake Plans

A `drake` plan is essentially a `tibble` that defines targets and the commands to build those targets. We can view `drake` plans by printing them to the console:

```{r}
plan # made from the previous example
```

One thing to be aware of is that the function `drake_plan()` does not evaluate its arguments, but only makes the plan. Thus, if you want any evaluation to occur within the construction of the plan, you will need some NSE magic.

In this example, the plan relies on `variable`, and so the result will change depending on the value of `variable` when the plan is run:

```{r}
variable <- "Species"
plan <- drake_plan(num_levels = nlevels(iris[, variable]))

## would compute nlevels on the "Sepal.Length" variable
# variable <- "Sepal.Length"
# make(plan)
```

Here, we ask that `variable` be evaluated when building the plan - this locks in the column setting as `"Species"`:

```{r}
variable <- "Species"
drake_plan(num_levels = nlevels(iris[, !!variable]))

## does not require variable to be in the current environment
# rm(variable)
# make(plan)
```

### Running Drake Plans

Drake plans are run by calling `make()`. This does several things. First it checks the **cache** to see if any targets need to be re-built, and then it proceeds to build all the targets, in some order that accounts for the dependencies between targets. (e.g. an analysis target that depends on a dataset target to be processed)

The manual has more information about how [Drake stores its cache](https://ropenscilabs.github.io/drake-manual/store.html#cache-formats) and how [Drake decides to rebuild targets](https://ropenscilabs.github.io/drake-manual/triggers.html).

Note that if there are file inputs, it is important that they are declared explicitly using e.g. `file_in()`, `knitr_in()`, and `file_out()`. This enables Drake to check if those files are changed and to rebuild targets that depend on the files if needed. Otherwise Drake will treat them as fixed strings.

```{r, eval = FALSE}
plan <- drake_plan(data = read.csv("some_data.csv"))
make(plan)

# make some changes to `some_data.csv`
make(plan) # will NOT rebuild the `data` target
```

```{r, eval = FALSE}
plan <- drake_plan(data = read.csv(file_in("some_data.csv")))
make(plan)

# make some changes to `some_data.csv`
make(plan) # will rebuild the `data` target
```
